Psychophysical description of musical elements in digital audio

### 2013-04-20 tex/phi+mus.tex
feito abstract

\documentclass[
 aip,
 jmp,
 amsmath,amssymb,
 reprint,
]{revtex4-1}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{multirow}
\usepackage{color}
\begin{document}
\preprint{AIP/123-QED}

\title{A quantitative approach to evolution of music
  and philosophy}

\author{Vilson Vieira}
 \homepage{http://automata.cc}
 \email{vilson@void.cc}
  \affiliation{ 
Instituto de F\'isica de S\~ao Carlos, Universidade de S\~ao Paulo (IFSC/USP)
}

\author{Renato Fabbri}
 \homepage{http://www.estudiolivre.org/el-user.php?view\_user=gk}
 \email{renato.fabbri@gmail.com}
 \altaffiliation[Also at ]{IFSC-USP}

\author{Gonzalo Travieso}
  \email{gonzalo@ifsc.usp.br}
 \altaffiliation[Also at ]{IFSC-USP}

@@

\date{\today}
\begin{abstract}
The development of new statistical and computational methods is
increasingly making it possible to bridge the gap between hard
sciences and humanities. In this study, we propose an approach based
on a quantitative evaluation of attributes of objects in fields of
humanities, from which concepts such as dialectics and opposition are
formally defined mathematically. As case studies, we analyzed the
temporal evolution of classical music and philosophy by obtaining data
for 8 features characterizing the corresponding fields for 7
well-known composers and philosophers, which were treated with
multivariate statistics and pattern recognition methods. A bootstrap
method was applied to avoid statistical bias caused by the small
sample data set, with which hundreds of artificial composers and
philosophers were generated, influenced by the 7 names originally
chosen.  Upon defining indices for opposition, skewness and
counter-dialectics, we confirmed the intuitive analysis of historians
in that classical music evolved according to a master-apprentice
tradition, while in philosophy changes were driven by
opposition. Though these case studies were meant only to show the
possibility of treating phenomena in humanities quantitatively,
including a quantitative measure of concepts such as dialectics and
opposition the results are encouraging for further application of the
approach presented here to many other areas, since it is entirely
generic.
\end{abstract}
\pacs{89.75.Fb,05.65.+b} % PACS, the Physics and Astronomy
\keywords{arts, music, musicology, philosophy, pattern recognition, statistics}
\maketitle

@@

Philosophy and natural sciences were born together in the 6th century
B.C. within Greek civilization~\cite{Russel}. Logic started with
Aristotle, geometry with Thales and Euclid, arithmetic with
Diophantus, while others created algebra, astronomy and politics, with
all fields being part of a common knowledge
space~\cite{Papineau}. What we know today as modern philosophy and
science were not originally separate, and arts were also present
almost universally in this space, not only as cultural expression. The
harmonic properties of music were of interest to philosophers such as
Pythagoras. Aristotle with his Poetics discussed a dramatic theory for
theatre. In this multidisciplinary space, music was addressed with
mathematics and science with philosophy. Quantitative methods were
applied to explain humanities, while the scientific method had its
origin in philosophy. The segregation of this space into philosophy,
science and arts was inevitable in the scientific revolution, thus
leading to its division into individual, independent areas. Although
specific, these areas are growing in complexity and their domains
overlap. Methods from specific areas are not sufficient to deal with
their complexity.

In this study, we cross the interdisciplinary borders and apply a
quantitative method to philosophy and music, in an attempt to
understand how humanities evolve. While philosophy and music have
their history constantly analyzed and discussed by critics, the nature
of their works --- text documents and music scores --- are difficult
to analyze and interpretation is always subjective. Here we propose a
generic approach to analyze features from any field in a quantitative
manner. The approach, described in the next Section, was then applied
to characterize composers of classical music and philosophers. The
data collected from scores assigned to these composers and
philosophers were treated with statistical and pattern recognition
methods, with the conclusions drawn being then compared to the
literature based on critics of music and philosophy.

>>>>

\documentclass[
 aip,
 jmp,
 amsmath,amssymb,
 reprint,
]{revtex4-1}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{multirow}
\usepackage{color}

\newcommand{\massa}{{\large \textsc{massa}}}
\newcommand{\mass}{{\large \textsc{mass}}}
\newcommand{\figgus}{{\large \textsc{figgus}}}

\begin{document}
\preprint{AIP/123-QED}

\title{Psychophysical description of musical elements in digital audio}

\author{Renato Fabbri}
 \homepage{http://www.estudiolivre.org/el-user.php?view\_user=gk}
 \email{renato.fabbri@gmail.com}
  \affiliation{ 
Instituto de F\'isica de S\~ao Carlos, Universidade de S\~ao Paulo (IFSC/USP)
}

@@

\date{\today}
\begin{abstract}

    The representation of the basic elements of music - such as notes, ornaments and intervalar structures - in terms of discrete audio signal is often used in software for music creation and design. Nevertheless, there is no unified approach that relates these elements to the sound discrete samples. In this article, each musical element is described by equations that represent the sonic time samples, which are all implemented in scripts within an open source software toolbox, referred to as \massa\ (Music and Audio in Sequences and Samples). The fundamental element, the musical note with duration, volume, pitch and timbre, is related quantitatively to the characteristics of the discrete-time signal. Internal variations, such as tremolos, vibratos and spectral fluctuations, are also considered, which enables the synthesis of notes inspired by real instruments and new sonorities. With this representation of notes, resources are provided for the generation of musical structures, such as rhythmic meter, pitch intervals and cycles. The efficacy of the natural representations was confirmed by the synthesis of small musical pieces using basic notes, incremented notes and notes in music. It is possible to synthesize whole albums through collage of the scripts and parameterization. 
    The sample-based analitical description, and the paradigm of open source implementation, enables cientific experiments in precise and trustful ways. In fact, \massa\ has already been employed by external users for diverse purposes. From these, it is mentioned a not identified in literature acoustic effect recognized by diverse individuals in mailing lists and uses related do art and education.

\end{abstract}
\pacs{*43.66.-x,43.66.+y,05.10.-a} % PACS, the Physics and Astronomy
\keywords{music, digital audio, statistics, psychophysics}
\maketitle

@@
